hbm_paper_results

### 2024-05-16 paper_lbcc_results.Rmd
looking at change in age effect on variance - plots

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam, ggeffects, effectsize)
#Sys.setenv("MC_CORES"=10L)
options("mc.cores")

#source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
#source("./cubic_scripts/R_scripts/ranked_welch_tests.R")

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam, ggeffects, effectsize, gratia)
#Sys.setenv("MC_CORES"=10L)
options("mc.cores")

source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
#source("./cubic_scripts/R_scripts/ranked_welch_tests.R")

@@

Cohen's F squared for study effects

```{r}
gam_centile_f2 <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_cohenfsq.csv"))
gam_centile_f2$combat <- "ComBat-GAM"

gamlss_centile_f2 <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_cohenfsq.csv"))
gamlss_centile_f2$combat <- "ComBatLS"

f2.df <- rbind(gam_centile_f2, gamlss_centile_f2)
f2.df <- f2.df %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

ggplot(f2.df) +
  geom_col(aes(x=pheno, y=fsq, fill=combat), position="dodge") + 
  facet_wrap(~pheno_cat, scales="free") +
  theme(axis.text.x = element_blank())
```

```{r}
f2.df.sum <- f2.df %>%
  group_by(pheno) %>%
  arrange(combat) %>%
  mutate(F2_diff = fsq - fsq[1]) #gamlss - gam value, + number means site effect is larger in combatLS

ggplot(f2.df.sum) +
  geom_col(aes(x=pheno, y=F2_diff), position="dodge") + 
  facet_wrap(~pheno_cat, scales="free") +
  theme(axis.text.x = element_blank())

f2.df.sum %>%
  group_by(pheno_cat) %>%
  summarise(mean_fsq_diff = mean(F2_diff)) %>%
  ggplot() + 
  geom_col(aes(x=pheno_cat, y=mean_fsq_diff), position="dodge")

f2.df.sum %>%
  #group_by(pheno_cat) %>%
  mutate(smaller_eff = case_when(F2_diff < 0 ~ "ComBatGAM",
                                    F2_diff > 0 ~ "ComBatLS")) %>%
  ggplot() +
  geom_bar(aes(x=smaller_eff, fill=pheno_cat), position="dodge")
```

## Centile Fans
**try plotting shapes of trajectores w/ and w/o study term to see how much they change the effect of age!

load unharmonized data
```{r}
raw.data <- fread(file="/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/lifespan_CN_imp-sites_euler.csv")
gmv.raw.mod <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler_raw_no.tbv_mod.rds"))

gmv.raw.pred.df <- pred.across.ages.lbcc(gmv.raw.mod, raw.data)

gmv.raw.pred.df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=f.sigma), color="red") +
  geom_line(aes(x=age_days, y=m.sigma), color="blue")

gmv.raw.mod.site <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler-raw_site.est_mod.rds"))

gmv.raw.pred.site.df <- pred.across.ages.lbcc.study(gmv.raw.mod.site, raw.data)

gmv.raw.pred.site.df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=f.sigma), color="red") +
  geom_line(aes(x=age_days, y=m.sigma), color="blue")

gmv.raw.pred.site.df <- gmv.raw.pred.site.df %>%
  mutate(study_incl = TRUE,
         combat= "raw")
gmv.raw.pred.df <- gmv.raw.pred.df %>%
  mutate(study_incl = FALSE,
         combat= "raw")
```

load ComBat-GAM-harmonized data
```{r}
cf.gam.data <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler_log-cf_cf.gam_batch.study_data.csv"))
gmv.cf.gam.mod <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_no.tbv_mod.rds"))

gmv.cf.gam.pred.df <- pred.across.ages.lbcc(gmv.cf.gam.mod, cf.gam.data)

gmv.cf.gam.pred.df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=f.sigma), color="red") +
  geom_line(aes(x=age_days, y=m.sigma), color="blue")

gmv.cf.gam.mod.site <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_site.est_mod.rds"))

gmv.cf.gam.pred.site.df <- pred.across.ages.lbcc.study(gmv.cf.gam.mod.site, cf.gam.data)

gmv.cf.gam.pred.site.df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=f.sigma), color="red") +
  geom_line(aes(x=age_days, y=m.sigma), color="blue")
```

```{r}
gmv.cf.gam.pred.site.df <- gmv.cf.gam.pred.site.df %>%
  mutate(study_incl = TRUE,
         combat= "combatGAM")
gmv.cf.gam.pred.df <- gmv.cf.gam.pred.df %>%
  mutate(study_incl = FALSE,
         combat= "combatGAM")

gmv.cf.gam.pred.df_full <- full_join(gmv.cf.gam.pred.site.df, gmv.cf.gam.pred.df)

gmv.cf.gam.pred.df_full %>%
  ggplot() +
  geom_line(aes(x=age_days, y=f.sigma, color=study_incl))
```

load ComBatLS-harmonized data
```{r}
cf.gamlss.data <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler_log-cf_cf.gamlss_batch.study_data.csv"))
gmv.cf.gamlss.mod <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_no.tbv_mod.rds"))

gmv.cf.gamlss.pred.df <- pred.across.ages.lbcc(gmv.cf.gamlss.mod, cf.gamlss.data)

gmv.cf.gamlss.pred.df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=f.sigma), color="red") +
  geom_line(aes(x=age_days, y=m.sigma), color="blue")

gmv.cf.gamlss.mod.site <- readRDS(paste0(der_path, "/GMV_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_site.est_mod.rds"))

gmv.cf.gamlss.pred.site.df <- pred.across.ages.lbcc.study(gmv.cf.gamlss.mod.site, cf.gamlss.data)

gmv.cf.gamlss.pred.site.df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=f.sigma), color="red") +
  geom_line(aes(x=age_days, y=m.sigma), color="blue")
```

```{r}
gmv.cf.gamlss.pred.site.df <- gmv.cf.gamlss.pred.site.df %>%
  mutate(study_incl = TRUE,
         combat= "combatLS")
gmv.cf.gamlss.pred.df <- gmv.cf.gamlss.pred.df %>%
  mutate(study_incl = FALSE,
         combat= "combatLS")

gmv.cf.gamlss.pred.df_full <- full_join(gmv.cf.gamlss.pred.site.df, gmv.cf.gamlss.pred.df)

gmv.cf.gamlss.pred.df_full %>%
  ggplot() +
  geom_line(aes(x=age_days, y=f.sigma, color=study_incl))
```

```{r}
all.pred.df <- reduce(list(gmv.cf.gam.pred.df_full, gmv.cf.gamlss.pred.df_full, gmv.raw.pred.site.df, gmv.raw.pred.df), full_join)

all.pred.df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=f.sigma, linetype=study_incl, color=combat))

all.pred.df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=m.sigma, linetype=study_incl, color=combat)) +
  ggtitle("GMV")

raw.data %>%
  ggplot() +
  geom_histogram(aes(x=age_days, fill=sex), position="identity", alpha=0.5)
```

See WMV
```{r}
#raw
wmv.raw.mod <- readRDS(paste0(der_path, "/wmv_lifespan-CN-imp-sites-euler_raw_no.tbv_mod.rds"))
wmv.raw.pred.df <- pred.across.ages.lbcc(wmv.raw.mod, raw.data)
wmv.raw.mod.site <- readRDS(paste0(der_path, "/wmv_lifespan-CN-imp-sites-euler-raw_site.est_mod.rds"))
wmv.raw.pred.site.df <- pred.across.ages.lbcc.study(wmv.raw.mod.site, raw.data)

wmv.raw.pred.site.df <- wmv.raw.pred.site.df %>%
  mutate(study_incl = TRUE,
         combat= "raw")
wmv.raw.pred.df <- wmv.raw.pred.df %>%
  mutate(study_incl = FALSE,
         combat= "raw")

#combat-gam
wmv.cf.gam.mod <- readRDS(paste0(der_path, "/wmv_lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_no.tbv_mod.rds"))
wmv.cf.gam.pred.df <- pred.across.ages.lbcc(wmv.cf.gam.mod, cf.gam.data)

wmv.cf.gam.mod.site <- readRDS(paste0(der_path, "/wmv_lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_site.est_mod.rds"))
wmv.cf.gam.pred.site.df <- pred.across.ages.lbcc.study(wmv.cf.gam.mod.site, cf.gam.data)

wmv.cf.gam.pred.site.df <- wmv.cf.gam.pred.site.df %>%
  mutate(study_incl = TRUE,
         combat= "combatGAM")
wmv.cf.gam.pred.df <- wmv.cf.gam.pred.df %>%
  mutate(study_incl = FALSE,
         combat= "combatGAM")

#combatLS
wmv.cf.gamlss.mod <- readRDS(paste0(der_path, "/wmv_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_no.tbv_mod.rds"))
wmv.cf.gamlss.pred.df <- pred.across.ages.lbcc(wmv.cf.gamlss.mod, cf.gamlss.data)

wmv.cf.gamlss.mod.site <- readRDS(paste0(der_path, "/wmv_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_site.est_mod.rds"))
wmv.cf.gamlss.pred.site.df <- pred.across.ages.lbcc.study(wmv.cf.gamlss.mod.site, cf.gamlss.data)

wmv.cf.gamlss.pred.site.df <- wmv.cf.gamlss.pred.site.df %>%
  mutate(study_incl = TRUE,
         combat= "combatLS")
wmv.cf.gamlss.pred.df <- wmv.cf.gamlss.pred.df %>%
  mutate(study_incl = FALSE,
         combat= "combatLS")
```

```{r}
wmv.all.pred.df <- reduce(list(wmv.raw.pred.site.df, wmv.raw.pred.df, wmv.cf.gam.pred.site.df, wmv.cf.gam.pred.df, wmv.cf.gamlss.pred.site.df, wmv.cf.gamlss.pred.df), full_join)

wmv.all.pred.df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=f.sigma, linetype=study_incl, color=combat))

wmv.all.pred.df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=m.sigma, linetype=study_incl, color=combat)) +
  ggtitle("WMV")

raw.data %>%
  ggplot() +
  geom_histogram(aes(x=age_days, fill=sex), position="identity", alpha=0.5)
```

See sGMV
```{r}
#raw
sgmv.raw.mod <- readRDS(paste0(der_path, "/sgmv_lifespan-CN-imp-sites-euler_raw_no.tbv_mod.rds"))
sgmv.raw.pred.df <- pred.across.ages.lbcc(sgmv.raw.mod, raw.data)
sgmv.raw.mod.site <- readRDS(paste0(der_path, "/sgmv_lifespan-CN-imp-sites-euler-raw_site.est_mod.rds"))
sgmv.raw.pred.site.df <- pred.across.ages.lbcc.study(sgmv.raw.mod.site, raw.data)

sgmv.raw.pred.site.df <- sgmv.raw.pred.site.df %>%
  mutate(study_incl = TRUE,
         combat= "raw")
sgmv.raw.pred.df <- sgmv.raw.pred.df %>%
  mutate(study_incl = FALSE,
         combat= "raw")

#combat-gam
sgmv.cf.gam.mod <- readRDS(paste0(der_path, "/sgmv_lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_no.tbv_mod.rds"))
sgmv.cf.gam.pred.df <- pred.across.ages.lbcc(sgmv.cf.gam.mod, cf.gam.data)

sgmv.cf.gam.mod.site <- readRDS(paste0(der_path, "/sgmv_lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_site.est_mod.rds"))
sgmv.cf.gam.pred.site.df <- pred.across.ages.lbcc.study(sgmv.cf.gam.mod.site, cf.gam.data)

sgmv.cf.gam.pred.site.df <- sgmv.cf.gam.pred.site.df %>%
  mutate(study_incl = TRUE,
         combat= "combatGAM")
sgmv.cf.gam.pred.df <- sgmv.cf.gam.pred.df %>%
  mutate(study_incl = FALSE,
         combat= "combatGAM")

#combatLS
sgmv.cf.gamlss.mod <- readRDS(paste0(der_path, "/sgmv_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_no.tbv_mod.rds"))
sgmv.cf.gamlss.pred.df <- pred.across.ages.lbcc(sgmv.cf.gamlss.mod, cf.gamlss.data)

sgmv.cf.gamlss.mod.site <- readRDS(paste0(der_path, "/sgmv_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_site.est_mod.rds"))
sgmv.cf.gamlss.pred.site.df <- pred.across.ages.lbcc.study(sgmv.cf.gamlss.mod.site, cf.gamlss.data)

sgmv.cf.gamlss.pred.site.df <- sgmv.cf.gamlss.pred.site.df %>%
  mutate(study_incl = TRUE,
         combat= "combatLS")
sgmv.cf.gamlss.pred.df <- sgmv.cf.gamlss.pred.df %>%
  mutate(study_incl = FALSE,
         combat= "combatLS")
```

```{r}
sgmv.all.pred.df <- reduce(list(sgmv.raw.pred.site.df, sgmv.raw.pred.df, sgmv.cf.gam.pred.site.df, sgmv.cf.gam.pred.df, sgmv.cf.gamlss.pred.site.df, sgmv.cf.gamlss.pred.df), full_join)

sgmv.all.pred.df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=f.sigma, linetype=study_incl, color=combat))

sgmv.all.pred.df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=m.sigma, linetype=study_incl, color=combat)) +
  ggtitle("sGMV")

raw.data %>%
  ggplot() +
  geom_histogram(aes(x=age_days, fill=sex), position="identity", alpha=0.5)
```

See Ventricles
```{r}
#raw
ven.raw.mod <- readRDS(paste0(der_path, "/Ventricles_lifespan-CN-imp-sites-euler_raw_no.tbv_mod.rds"))
ven.raw.pred.df <- pred.across.ages.lbcc(ven.raw.mod, raw.data)
# ven.raw.mod.site <- readRDS(paste0(der_path, "/Ventricles_lifespan-CN-imp-sites-euler_raw_site.est_mod.rds"))
# ven.raw.pred.site.df <- pred.across.ages.lbcc.study(ven.raw.mod.site, raw.data)

# ven.raw.pred.site.df <- ven.raw.pred.site.df %>%
#   mutate(study_incl = TRUE,
#          combat= "raw")
ven.raw.pred.df <- ven.raw.pred.df %>%
  mutate(study_incl = FALSE,
         combat= "raw")

#combat-gam
ven.cf.gam.mod <- readRDS(paste0(der_path, "/Ventricles_lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_no.tbv_mod.rds"))
ven.cf.gam.pred.df <- pred.across.ages.lbcc(ven.cf.gam.mod, cf.gam.data)

ven.cf.gam.mod.site <- readRDS(paste0(der_path, "/Ventricles_lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_site.est_mod.rds"))
ven.cf.gam.pred.site.df <- pred.across.ages.lbcc.study(ven.cf.gam.mod.site, cf.gam.data)

ven.cf.gam.pred.site.df <- ven.cf.gam.pred.site.df %>%
  mutate(study_incl = TRUE,
         combat= "combatGAM")
ven.cf.gam.pred.df <- ven.cf.gam.pred.df %>%
  mutate(study_incl = FALSE,
         combat= "combatGAM")

#combatLS
ven.cf.gamlss.mod <- readRDS(paste0(der_path, "/Ventricles_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_no.tbv_mod.rds"))
ven.cf.gamlss.pred.df <- pred.across.ages.lbcc(ven.cf.gamlss.mod, cf.gamlss.data)

ven.cf.gamlss.mod.site <- readRDS(paste0(der_path, "/Ventricles_lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_site.est_mod.rds"))
ven.cf.gamlss.pred.site.df <- pred.across.ages.lbcc.study(ven.cf.gamlss.mod.site, cf.gamlss.data)

ven.cf.gamlss.pred.site.df <- ven.cf.gamlss.pred.site.df %>%
  mutate(study_incl = TRUE,
         combat= "combatLS")
ven.cf.gamlss.pred.df <- ven.cf.gamlss.pred.df %>%
  mutate(study_incl = FALSE,
         combat= "combatLS")
```

```{r}
ven.all.pred.df <- reduce(list(ven.raw.pred.df, ven.cf.gam.pred.site.df, ven.cf.gam.pred.df, ven.cf.gamlss.pred.site.df, ven.cf.gamlss.pred.df), full_join) #ven.raw.pred.site.df

ven.all.pred.df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=f.sigma, linetype=study_incl, color=combat))

ven.all.pred.df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=m.sigma, linetype=study_incl, color=combat)) +
  ggtitle("Ventricles")

raw.data %>%
  ggplot() +
  geom_histogram(aes(x=age_days, fill=sex), position="identity", alpha=0.5) +
  ggtitle("Sample distribution")
```
